---
title: "Anderson-Darling Test"
author: "Stefan Kloppenborg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
csl: ieee-with-url.csl
references:
- id: CMH-17-1G
  type: report
  number: CMH-17-1G
  title: Composites Materials Handbook, Volume 1. Polymer Matrix Composites Guideline for Characterization of Structural Materials
  issued:
    year: 2012
    month: 03
  publisher: SAE International
- id: Stephens1974
  type: article
  author: 
  - given: M.A.
    family: Stephens
  title: EDF Statistics for Goodness of Fit and Some Comparisons
  container-title: Journal of the American Statistical Association
  volume: 69
  issue: 347
  page: 730-737
  issued:
    year: 1974
    month: 09
- id: Marsaglia2004
  type: article
  author:
  - given: George
    family: Marsaglia
  - given: John
    family: Marsaglia
  title: Evaluating the Anderson-Darling Distribution
  container-title: Journal of Statistical Software
  volume: 9
  issue: 2
  issued:
    year: 2004
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(tidyverse)
library(cmstatr)
```

# Comparison of P-Value for Known and Unknown Parameters
There are various methods published for calculating the p-value given an Anderson-Darling statistic. [@Stephens1974] provides Anderson-Darling statistics at various significance levels, and [@Marsaglia2004] provides an algorithm for doing so. However, [@CMH-17-1G] provides a formula for the OSL, or Observed Significance Level (which is the same as a p-value). At first glance, one might expect these to be the same, but if one actually computes both for a given Anderson-Darling statistic, one will find that they differ. The reason for this difference are the assumptions. The p-vales given by [@Stephens1974] assume that the parameters of the distribution are known (i.e. for a normal distribution, that the population mean and population standard deviation are known). The p-value given by [@CMH-17-1G] assume that the parameters of the distribution are unknown and that the parameters used are simply estimates based on the sample.

Both the method in [@Marsaglia2004] and [@CMH-17-1G] have a correction for sample size. We'll plot the OSL and the p-value for different values of $AD$ (the Anderson-Darling statistic) and a fixed value of $n$. For now, we'll choose a small value of $n$ and a large value, say $n=5$ and $n=100$. Both methods are provided in the `cmstatr` package and we will use that implementation.

```{r}
expand.grid(
  n = c(5, 100),
  AD = c(0, 10 ^ (seq(-1.1, 0.5, length.out = 100)))
) %>%
  rowwise() %>%
  mutate(
    p_known_param = ad_p_known_param(AD, n),
    p_unknown_param = ad_p_unknown_param(AD, n)
  ) %>%
  gather(stat, value, p_known_param, p_unknown_param) %>%
  group_by(n, stat) %>%
  ggplot(aes(x = AD, y = value, color = stat)) +
  geom_line() +
  facet_grid(n ~ .)
```

Both the p-value and the OSL take values of $1.0$ when $AD=0$ and both tend towards $0$ as $AD$ increases. However, the shape of both is quite different. This is due to the uncertainty in the parameters for the distribution to which the data is being compared with.

# References
