---
title: "cmstatr Validation"
author: "Stefan Kloppenborg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cmstatr Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
csl: ieee.csl
references:
- id: "ASAP2008"
  type: "report"
  number: "ASAP-2008"
  author:
  - given: K.S.
    family: Raju
  - given: J.S.
    family: Tomblin
  title: "AGATE Statistical Analysis Program"
  issued:
    year: 2008
  publisher: Wichita State University
- id: Stephens1987
  type: article
  author:
    - given: F.W.
      family: Scholz
    - given: M.A,
      family: Stephens
  title: K-Sample Anderson-Darling Tests
  container-title: Journal of the American Statistical Association
  volume: "82"
  issue: "399"
  page: 918-924
  issued:
    year: "1987"
    month: "09"
  DOI: 10.1080/01621459.1987.10478517
  URL: https://doi.org/10.1080/01621459.1987.10478517
- id: "CMH-17-1G"
  type: report
  number: "CMH-17-1G"
  title: Composite Materials Handbook, Volume 1. Polymer Matrix Composites
    Guideline for Characterization of Structural Materials
  publisher: SAE International
  issued:
    year: "2012"
    month: "03"
- id: "STAT-17"
  type: report
  number: "STAT-17 Rev 5"
  author:
    - literal: Materials Sciences Corporation
  title: CMH-17 Statistical Analysis for B-Basis and A-Basis Values
  publisher: Materials Sciences Corporation
  publisher-place: Horsham, PA
  issued:
    year: "2008"
    month: "01"
    day: "08"
- id: Vangel1994
  type: article
  author:
    - given: Mark
      family: Vangel
  title: One-Sided Nonparametric Tolerance Limits
  container-title: Communications in Statistics - Simulation and Computation
  volume: "23"
  issue: "4"
  page: 1137-1154
  issued:
    year: "1994"
  DOI: 10.1080/03610919408813222

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# 1. Introduction

This vignette is intended to contain the same validation that is included
in the test suite within the `cmstatr` package, but in a format that is
easier for a human to read. The intent is that this vignette will include
only those validations that are included in the test suite, but that the
test suite may include more tests than are shown in this vignette.

The following packages will be used in this validation. The version of
each package used is listed at the end of this vignette.

```{r message=FALSE, warning=FALSE}
library(cmstatr)
library(dplyr)
library(testthat)
```

Throughout this vignette, the `testthat` package will be used. Expressions
such as `expect_equal` are used to ensure that the two values are equal
(within some tolerance). If this expectation is not true, the vignette will
fail to build. As an example, the following expression checks that the value
`10` is equal to `10.1` within a tolerance of `0.5`. Such an expectation
should be satisfied.


```{r}
expect_equal(10, 10.1, tolerance = 0.5)
```

# 2. Validation Table
The following table provides a cross-reference between the various
functions of the `cmstatr` package and the tests shown within this
vignette. The sections in this vignette are organized by data set.
Not all checks are performed on all data sets.


Function                        | Tests
--------------------------------|---------------------------
`ad_ksample()`                  | [Section 3.1](#cf-ad)
`anderson_darling_normal()`     | 
`anderson_darling_lognormal()`  | 
`anderson_darling_weibull()`    | 
`basis_normal()`                |
`basis_lognormal()`             |
`basis_weibull()`               |
`basis_pooled_cv()`             |
`basis_pooled_sd()`             |
`basis_hk_ext()`                |
`basis_nonpara_large_sample()`  |
`basis_anova()`                 |
`calc_cv_star()`                |
`cv()`                          |
`equiv_change_mean()`           |
`equiv_mean_extremum()`         |
`hk_ext_z()`                    |
`hk_ext_z_j_opt()`              |
`k_equiv()`                     |
`k_factor_normal()`             |
`levene_test()`                 |
`maximum_normed_residual()`     |
`nonpara_binomial_rank()`       |
`normalize_group_mean()`        |
`normalize_ply_thickness()`     |
`transform_mod_cv_ad()`         |
`transform_mod_cv()`            |


# 3. `carbon.fabric` Data Set

This data set is example data that is provided with `cmstatr`.
The first few rows of this data are shown below.

```{r}
head(carbon.fabric)
```


## 3.1. Anderson--Darling k-Sample Test {#cf-ad}

This data was entered into ASAP 2008 [@ASAP2008] and the reported
Anderson--Darling k--Sample test statistics were recorded, as were
the conclusions.

The value of the test statistic reported by `cmstatr` and that reported
by ASAP 2008 differ by a factor of $k - 1$, as do the critical values
used. As such, the conclusion of the tests are identical. This is
described in more detail in the
[Anderson--Darling k--Sample Vignette](adktest1.html).

When the RTD warp-tension data from this data set is entered into ASAP 2008,
it reports a test statistic of 0.456 and fails to reject the null hypothesis
that the batches are drawn from the same distribution. Adjusting for the
different definition of the test statistic, the results given by `cmstatr`
are very similar


```{r}
res <- carbon.fabric %>%
  filter(test == "WT") %>%
  filter(condition == "RTD") %>%
  ad_ksample(strength, batch)

expect_equal(res$ad / (res$k - 1), 0.456, tolerance = 0.002)
expect_false(res$reject_same_dist)

res
```

When the ETW warp-tension data from this data set are entered into ASAP 2008,
the reported test statistic is 1.604 and it fails to reject the null
hypothesis that the batches are drawn from the same distribution. Adjusting
for the different definition of the test statistic, `cmstatr` gives nearly
identical results.

```{r}
res <- carbon.fabric %>%
  filter(test == "WT") %>%
  filter(condition == "ETW") %>%
  ad_ksample(strength, batch)

expect_equal(res$ad / (res$k - 1), 1.604, tolerance = 0.002)
expect_false(res$reject_same_dist)

res
```


# X. Comparison with Examples from CMH-17-1G
## X.X Section 8.3.11.1.1

[@CMH-17-1G] provides an example data set and results from
ASAP [@ASAP2008] and STAT17 [@STAT-17]. This example data set is duplicated
below:

```{r}
dat_8_3_11_1_1 <- tribble(
  ~batch, ~strength, ~condition,
  1, 118.3774604, "CTD", 1, 84.9581364, "RTD", 1, 83.7436035, "ETD",
  1, 123.6035612, "CTD", 1, 92.4891822, "RTD", 1, 84.3831677, "ETD",
  1, 115.2238092, "CTD", 1, 96.8212659, "RTD", 1, 94.8030433, "ETD",
  1, 112.6379744, "CTD", 1, 109.030325, "RTD", 1, 94.3931537, "ETD",
  1, 116.5564277, "CTD", 1, 97.8212659, "RTD", 1, 101.702222, "ETD",
  1, 123.1649896, "CTD", 1, 100.921519, "RTD", 1, 86.5372121, "ETD",
  2, 128.5589027, "CTD", 1, 103.699444, "RTD", 1, 92.3772684, "ETD",
  2, 113.1462103, "CTD", 2, 93.7908212, "RTD", 2, 89.2084024, "ETD",
  2, 121.4248107, "CTD", 2, 107.526709, "RTD", 2, 100.686001, "ETD",
  2, 134.3241906, "CTD", 2, 94.5769704, "RTD", 2, 81.0444192, "ETD",
  2, 129.6405117, "CTD", 2, 93.8831373, "RTD", 2, 91.3398070, "ETD",
  2, 117.9818658, "CTD", 2, 98.2296605, "RTD", 2, 93.1441939, "ETD",
  3, 115.4505226, "CTD", 2, 111.346590, "RTD", 2, 85.8204168, "ETD",
  3, 120.0369467, "CTD", 2, 100.817538, "RTD", 3, 94.8966273, "ETD",
  3, 117.1631088, "CTD", 3, 100.382203, "RTD", 3, 95.8068520, "ETD",
  3, 112.9302797, "CTD", 3, 91.5037811, "RTD", 3, 86.7842252, "ETD",
  3, 117.9114501, "CTD", 3, 100.083233, "RTD", 3, 94.4011973, "ETD",
  3, 120.1900159, "CTD", 3, 95.6393615, "RTD", 3, 96.7231171, "ETD",
  3, 110.7295966, "CTD", 3, 109.304779, "RTD", 3, 89.9010384, "ETD",
  3, 100.078562, "RTD", 3, 99.1205847, "RTD", 3, 89.3672306, "ETD",
  1, 106.357525, "ETW", 1, 99.0239966, "ETW2",
  1, 105.898733, "ETW", 1, 103.341238, "ETW2",
  1, 88.4640082, "ETW", 1, 100.302130, "ETW2",
  1, 103.901744, "ETW", 1, 98.4634133, "ETW2",
  1, 80.2058219, "ETW", 1, 92.2647280, "ETW2",
  1, 109.199597, "ETW", 1, 103.487693, "ETW2",
  1, 61.0139431, "ETW", 1, 113.734763, "ETW2",
  2, 99.3207107, "ETW", 2, 108.172659, "ETW2",
  2, 115.861770, "ETW", 2, 108.426732, "ETW2",
  2, 82.6133082, "ETW", 2, 116.260375, "ETW2",
  2, 85.3690411, "ETW", 2, 121.049610, "ETW2",
  2, 115.801622, "ETW", 2, 111.223082, "ETW2",
  2, 44.3217741, "ETW", 2, 104.574843, "ETW2",
  2, 117.328077, "ETW", 2, 103.222552, "ETW2",
  2, 88.6782903, "ETW", 3, 99.3918538, "ETW2",
  3, 107.676986, "ETW", 3, 87.3421658, "ETW2",
  3, 108.960241, "ETW", 3, 102.730741, "ETW2",
  3, 116.122640, "ETW", 3, 96.3694916, "ETW2",
  3, 80.2334815, "ETW", 3, 99.5946088, "ETW2",
  3, 106.145570, "ETW", 3, 97.0712407, "ETW2",
  3, 104.667866, "ETW",
  3, 104.234953, "ETW"
)
dat_8_3_11_1_1
```

### X.X.X Maximum Normed Residual Test
CMH-17-1G Table 8.3.11.1.1(a) provides results of the MNR test from
ASAP for this data set. Batches 2 and 3 of the ETW data is considered
here and the results of `cmstatr` are compared with those published in
CMH-17-1G.

For Batch 2 of the ETW data, the results match those published in the
handbook within a small tolerance.

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW" & batch == 2) %>%
  maximum_normed_residual(strength, alpha = 0.05)

expect_equal(res$mnr, 2.008, tolerance = 0.001)
expect_equal(res$crit, 2.127, tolerance = 0.001)
expect_equal(res$n_outliers, 0)

res
```

Similarly, for Batch 3 of the ETW data, the results of `cmstatr` match
the results published in the handbook within a small tolerance.

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW" & batch == 3) %>%
  maximum_normed_residual(strength, alpha = 0.05)

expect_equal(res$mnr, 2.119, tolerance = 0.001)
expect_equal(res$crit, 2.020, tolerance = 0.001)
expect_equal(res$n_outliers, 1)

res
```


### X.X.X Anderson--Darling k--Sample Test

For the ETW condition, the ADK test statistic given in @[CMH-17-1G] is
$ADK = 0.793$ and the test concludes that the samples come from the
same distribution. Noting that `cmstatr` uses the definition of the
test statistic given in [@Stephens1987], so the test statistic given
by `cmstatr` differs from that given by ASAP by a factor of $k - 1$,
as described in the [Anderson--Darling k--Sample Vignette](adktest1.html).

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW") %>%
  ad_ksample(strength, batch)

expect_equal(res$ad / (res$k - 1), 0.793, tolerance = 0.003)
expect_false(res$reject_same_dist)

res
```

Similarly, for the ETW2 condition, the test statistic given in [@CMH-17-1G]
is $ADK = 3.024$ and the test concludes that the samples come from different
distributions. This matches `cmstatr`

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW2") %>%
  ad_ksample(strength, batch)

expect_equal(res$ad / (res$k - 1), 3.024, tolerance = 0.003)
expect_true(res$reject_same_dist)

res
```

### X.X.X Anderson--Darling Tests for Distribution

CMH-17-1G Section 8.3.11.2.1 contains results from STAT17 for
the "observed significance level" from the Anderson--Darling test
for various distributions. In this section, the ETW condition from the
present data set is used. The published results are given in the 
following table. The results from `cmstatr` are below and are very
similar to those from STAT17.

Distribution | OSL
-------------|------------
Normal       | 0.006051
Lognormal    | 0.000307
Weibull      | 0.219

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW") %>%
  anderson_darling_normal(strength)
expect_equal(res$osl, 0.006051, tolerance = 1e-6)
res
```
```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW") %>%
  anderson_darling_lognormal(strength)
expect_equal(res$osl, 0.000307, tolerance = 1e-6)
res
```

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW") %>%
  anderson_darling_weibull(strength)
expect_equal(res$osl, 0.0219, tolerance = 1e-4)
res
```


### X.X.X Levene's Test (Between Conditions)

??? The test statistic is different in the handbook????

```{r}
dat_8_3_11_1_1 %>%
  filter(condition != "ETW" & condition != "ETW2") %>%
  levene_test(strength, condition)
```

### X.X.X Levene's Test (Between Batches)

CMH-17-1G Section 8.3.11.2.2 provides output from STAT17. The
ETW2 condition from the present data set was analyzed by STAT17
and that software reported an F statistic of 0.123 from Levene's
test when comparing the variance of the batches within this condition.
The result from `cmstatr` is similar.

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW2") %>%
  levene_test(strength, batch)
expect_equal(res$f, 0.123, tolerance = 1e-3)
res
```

Similarly, the published value of the F statistic for the CTD condition is 
$3.850$. `cmstatr` produces very similar results.

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "CTD") %>%
  levene_test(strength, batch)
expect_equal(res$f, 3.850, tolerance = 1e-3)
res
```



### X.X.X Nonparametric Basis Values
CMH-17-1G Section 8.3.11.2.1 provides STAT17 outputs for the ETW
condition of the present data set. The nonparametric Basis values are listed.
In this case, the Hanson--Koopmans method is used. The published A-Basis
value is 13.0 and the B-Basis is 37.9.


```{r}
dat_8_3_11_1_1 %>%
  filter(condition == "ETW") %>%
  basis_hk_ext(strength, method = "woodward-frawley", p = 0.99, conf = 0.95)
```

TODO: THIS IS NOT THE RESULT THAT IS PUBLISHED!!!!!!!

```{r}
dat_8_3_11_1_1 %>%
  filter(condition == "ETW") %>%
  basis_hk_ext(strength, method = "optimum-order", p = 0.90, conf = 0.95)
```



### X.X.X Single-Point ANOVA Basis Value
CMH-17-1G Section 8.3.11.2.2 provides output from STAT17 for
the ETW2 condition from the present data set. STAT17 reports
A- and B-Basis values based on the ANOVA method of 34.6 and
63.2, respectively. The results from `cmstatr` are similar.

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW2") %>%
  basis_anova(strength, batch, override = "number_of_groups",
              p = 0.99, conf = 0.95)
expect_equal(res$basis, 34.6, tolerance = 5e-3)
res
```

```{r}
res <- dat_8_3_11_1_1 %>%
  filter(condition == "ETW2") %>%
  basis_anova(strength, batch, override = "number_of_groups")
expect_equal(res$basis, 63.2, tolerance = 5e-3)
res
```


# Other Data Sets
This section contains various small data sets. In most cases, these data sets
were generated randomly for the purpose of comparing `cmstatr` to other
software.

## Anderson--Darling Test (Normal)
The following data set was randomly generated. When this is
entered into STAT17 [@STAT-17], that software gives the value
$OSL = 0.465$, which matches the result of `cmstatr`
within a small margin.

```{r}
dat <- data.frame(
  strength = c(
    137.4438,
    139.5395,
    150.89,
    141.4474,
    141.8203,
    151.8821,
    143.9245,
    132.9732,
    136.6419,
    138.1723,
    148.7668,
    143.283,
    143.5429,
    141.7023,
    137.4732,
    152.338,
    144.1589,
    128.5218
))
res <- anderson_darling_normal(dat, strength)

expect_equal(res$osl, 0.465, tolerance = 0.002)

res
```

## Anderson--Darling Test (Lognormal)
The following data set was randomly generated. When this is
entered into STAT17 [@STAT-17], that software gives the value
$OSL = 0.480$, which matches the result of `cmstatr`
within a small margin.

```{r}
dat <- data.frame(
  strength = c(
    137.4438,
    139.5395,
    150.89,
    141.4474,
    141.8203,
    151.8821,
    143.9245,
    132.9732,
    136.6419,
    138.1723,
    148.7668,
    143.283,
    143.5429,
    141.7023,
    137.4732,
    152.338,
    144.1589,
    128.5218
))

res <- anderson_darling_lognormal(dat, strength)

expect_equal(res$osl, 0.480, tolerance = 0.002)

res
```

## Anderson--Darling Test (Weibull)
The following data set was randomly generated. When this is
entered into STAT17 [@STAT-17], that software gives the value
$OSL = 0.179$, which matches the result of `cmstatr`
within a small margin.

```{r}
dat <- data.frame(
  strength = c(
    137.4438,
    139.5395,
    150.89,
    141.4474,
    141.8203,
    151.8821,
    143.9245,
    132.9732,
    136.6419,
    138.1723,
    148.7668,
    143.283,
    143.5429,
    141.7023,
    137.4732,
    152.338,
    144.1589,
    128.5218
))

res <- anderson_darling_weibull(dat, strength)

expect_equal(res$osl, 0.179, tolerance = 0.002)

res
```



## X.X Extended Hanson--Koopman B-Basis
The following random numbers were generated.

```{r}
dat <- c(
  139.6734,
  143.0032,
  130.4757,
  144.8327,
  138.7818,
  136.7693,
  148.636,
  131.0095,
  131.4933,
  142.8856,
  158.0198,
  145.2271,
  137.5991,
  139.8298,
  140.8557,
  137.6148,
  131.3614,
  152.7795,
  145.8792,
  152.9207,
  160.0989,
  145.192,
  128.6383,
  141.5992,
  122.5297,
  159.8209,
  151.672,
  159.0156
)
```

All of the numbers above were input into STAT17 and the reported B-Basis
value using the Optimum Order nonparametric method was 122.36798. This
result matches the results of `cmstatr` within a small margin.

```{r}
res <- basis_hk_ext(x = dat, p = 0.9, conf = 0.95,
                    method = "optimum-order")
expect_equal(res$basis, 122.36798, tolerance = 0.001)
res
```

The last two observations from the above data set were discarded, leaving
26 observations. This smaller data set was input into STAT17 and that
software calculated a B-Basis value of 121.57073 using the Optimum
Order nonparametric method. `cmstatr` reports a very similar number.

```{r}
res <- basis_hk_ext(x = head(dat, 26), p = 0.9, conf = 0.95,
                    method = "optimum-order")
expect_equal(res$basis, 121.57073, tolerance = 0.001)
res
```

The same data set was further reduced such that only the first 22
observations were included. This smaller data set was input into STAT17
and that
software calculated a B-Basis value of 128.82397 using the Optimum
Order nonparametric method. `cmstatr` reports a very similar number.

```{r}
res <- basis_hk_ext(x = head(dat, 22), p = 0.9, conf = 0.95,
                    method = "optimum-order")
expect_equal(res$basis, 128.82397, tolerance = 0.001)
res
```


# X. Comparison with Literature
In this section, results from `cmstatr` are compared with values published
in literature.

## X.X Anderson--Darling K--Sample Test
[@Stephens1987] provides example data that compares measurements obtained
in four labs. Their paper gives values of the ADK test statistic as well as
p-values.

The data in [@Stephens1987] is as follows:

```{r}
dat_ss1987 <- data.frame(
  smoothness = c(
    38.7, 41.5, 43.8, 44.5, 45.5, 46.0, 47.7, 58.0,
    39.2, 39.3, 39.7, 41.4, 41.8, 42.9, 43.3, 45.8,
    34.0, 35.0, 39.0, 40.0, 43.0, 43.0, 44.0, 45.0,
    34.0, 34.8, 34.8, 35.4, 37.2, 37.8, 41.2, 42.8
  ),
  lab = c(rep("A", 8), rep("B", 8), rep("C", 8), rep("D", 8))
)
dat_ss1987
```

[@Stephens1987] lists the corresponding test statistics
$A_{akN}^2 = 8.3926$ and $\sigma_N = 1.2038$ with the p-value
$p = 0.0022$. These match the result of `cmstatr` within a small margin.

```{r}
res <- ad_ksample(dat_ss1987, smoothness, lab)

expect_equal(res$ad, 8.3926, tolerance = 1e-3)
expect_equal(res$sigma, 1.2038, tolerance = 1e-3)
expect_equal(res$p, 0.00226, tolerance = 1e-3)

res
```


# Comparison with Published Factors
Various factors, such as tolerance limit factors, are published in various
publications. This section compares those published factors with those
computed by `cmstatr`.

## Normal kB Factors
B-Basis tolerance limit factors assuming a normal distribution are published in
CMH-17-1G. Those factors are reproduced below and are compared with the
results of `cmstatr`. The published factors and those computed by `cmstatr`
are quite similar.

```{r message=FALSE, warning=FALSE}
tribble(
  ~n, ~kB_published,
  2, 20.581, 36, 1.725, 70, 1.582, 104, 1.522,
  3, 6.157, 37, 1.718, 71, 1.579, 105, 1.521,
  4, 4.163, 38, 1.711, 72, 1.577, 106, 1.519,
  5, 3.408, 39, 1.704, 73, 1.575, 107, 1.518,
  6, 3.007, 40, 1.698, 74, 1.572, 108, 1.517,
  7, 2.756, 41, 1.692, 75, 1.570, 109, 1.516,
  8, 2.583, 42, 1.686, 76, 1.568, 110, 1.515,
  9, 2.454, 43, 1.680, 77, 1.566, 111, 1.513,
  10, 2.355, 44, 1.675, 78, 1.564, 112, 1.512,
  11, 2.276, 45, 1.669, 79, 1.562, 113, 1.511,
  12, 2.211, 46, 1.664, 80, 1.560, 114, 1.510,
  13, 2.156, 47, 1.660, 81, 1.558, 115, 1.509,
  14, 2.109, 48, 1.655, 82, 1.556, 116, 1.508,
  15, 2.069, 49, 1.650, 83, 1.554, 117, 1.507,
  16, 2.034, 50, 1.646, 84, 1.552, 118, 1.506,
  17, 2.002, 51, 1.642, 85, 1.551, 119, 1.505,
  18, 1.974, 52, 1.638, 86, 1.549, 120, 1.504,
  19, 1.949, 53, 1.634, 87, 1.547, 121, 1.503,
  20, 1.927, 54, 1.630, 88, 1.545, 122, 1.502,
  21, 1.906, 55, 1.626, 89, 1.544, 123, 1.501,
  22, 1.887, 56, 1.623, 90, 1.542, 124, 1.500,
  23, 1.870, 57, 1.619, 91, 1.540, 125, 1.499,
  24, 1.854, 58, 1.616, 92, 1.539, 126, 1.498,
  25, 1.839, 59, 1.613, 93, 1.537, 127, 1.497,
  26, 1.825, 60, 1.609, 94, 1.536, 128, 1.496,
  27, 1.812, 61, 1.606, 95, 1.534, 129, 1.495,
  28, 1.800, 62, 1.603, 96, 1.533, 130, 1.494,
  29, 1.789, 63, 1.600, 97, 1.531, 131, 1.493,
  30, 1.778, 64, 1.597, 98, 1.530, 132, 1.492,
  31, 1.768, 65, 1.595, 99, 1.529, 133, 1.492,
  32, 1.758, 66, 1.592, 100, 1.527, 134, 1.491,
  33, 1.749, 67, 1.589, 101, 1.526, 135, 1.490,
  34, 1.741, 68, 1.587, 102, 1.525, 136, 1.489,
  35, 1.733, 69, 1.584, 103, 1.523, 137, 1.488
) %>%
  arrange(n) %>%
  mutate(kB_cmstatr = k_factor_normal(n, p = 0.9, conf = 0.95)) %>%
  rowwise() %>%
  mutate(abs_difference = expect_less_than(
    abs(kB_published - kB_cmstatr), 5e-3))
```

## Normal kA Factors
A-Basis tolerance limit factors assuming a normal distribution are published in
CMH-17-1G. Those factors are reproduced below and are compared with the
results of `cmstatr`. The published factors and those computed by `cmstatr`
are quite similar.

```{r message=FALSE, warning=FALSE}
tribble(
  ~n, ~kA_published,
  2, 37.094, 36, 2.983, 70, 2.765, 104, 2.676,
  3, 10.553, 37, 2.972, 71, 2.762, 105, 2.674,
  4, 7.042, 38, 2.961, 72, 2.758, 106, 2.672,
  5, 5.741, 39, 2.951, 73, 2.755, 107, 2.671,
  6, 5.062, 40, 2.941, 74, 2.751, 108, 2.669,
  7, 4.642, 41, 2.932, 75, 2.748, 109, 2.667,
  8, 4.354, 42, 2.923, 76, 2.745, 110, 2.665,
  9, 4.143, 43, 2.914, 77, 2.742, 111, 2.663,
  10, 3.981, 44, 2.906, 78, 2.739, 112, 2.662,
  11, 3.852, 45, 2.898, 79, 2.736, 113, 2.660,
  12, 3.747, 46, 2.890, 80, 2.733, 114, 2.658,
  13, 3.659, 47, 2.883, 81, 2.730, 115, 2.657,
  14, 3.585, 48, 2.876, 82, 2.727, 116, 2.655,
  15, 3.520, 49, 2.869, 83, 2.724, 117, 2.654,
  16, 3.464, 50, 2.862, 84, 2.721, 118, 2.652,
  17, 3.414, 51, 2.856, 85, 2.719, 119, 2.651,
  18, 3.370, 52, 2.850, 86, 2.716, 120, 2.649,
  19, 3.331, 53, 2.844, 87, 2.714, 121, 2.648,
  20, 3.295, 54, 2.838, 88, 2.711, 122, 2.646,
  21, 3.263, 55, 2.833, 89, 2.709, 123, 2.645,
  22, 3.233, 56, 2.827, 90, 2.706, 124, 2.643,
  23, 3.206, 57, 2.822, 91, 2.704, 125, 2.642,
  24, 3.181, 58, 2.817, 92, 2.701, 126, 2.640,
  25, 3.158, 59, 2.812, 93, 2.699, 127, 2.639,
  26, 3.136, 60, 2.807, 94, 2.697, 128, 2.638,
  27, 3.116, 61, 2.802, 95, 2.695, 129, 2.636,
  28, 3.098, 62, 2.798, 96, 2.692, 130, 2.635,
  29, 3.080, 63, 2.793, 97, 2.690, 131, 2.634,
  30, 3.064, 64, 2.789, 98, 2.688, 132, 2.632,
  31, 3.048, 65, 2.785, 99, 2.686, 133, 2.631,
  32, 3.034, 66, 2.781, 100, 2.684, 134, 2.630,
  33, 3.020, 67, 2.777, 101, 2.682, 135, 2.628,
  34, 3.007, 68, 2.773, 102, 2.680, 136, 2.627,
  35, 2.995, 69, 2.769, 103, 2.678, 137, 2.626
) %>%
  arrange(n) %>%
  mutate(kA_cmstatr = k_factor_normal(n, p = 0.99, conf = 0.95)) %>%
  rowwise() %>%
  mutate(abs_difference = expect_less_than(
    abs(kA_published - kA_cmstatr), 5e-4))
```

## Nonparametric B-Basis Extended Hanson--Koopman
Vangel [@Vangel1994] provides extensive tables of $z$ for the case where
$i=1$ and $j$ is the median observation. This section checks the results of
`cmstatr`'s function against those tables. Only the odd values of $n$
are checked so that the median is a single observation. The unit tests for
the `cmstatr` package include checks of a variety of values of $p$ and
confidence, but only the factors for B-Basis are checked here.

```{r}
tribble(
  ~n, ~z,
  3,  28.820048,
  5,  6.1981307,
  7,  3.4780112,
  9,  2.5168762,
  11, 2.0312134,
  13, 1.7377374,
  15, 1.5403989,
  17, 1.3979806,
  19, 1.2899172,
  21, 1.2048089,
  23, 1.1358259,
  25, 1.0786237,
  27, 1.0303046,
) %>%
  rowwise() %>%
  mutate(
    z_calc = hk_ext_z(n, 1, ceiling(n / 2), p = 0.90, conf = 0.95)
  ) %>%
  mutate(abs_difference = expect_lt(abs(z - z_calc), 5e-5))
```

## Nonparametric A-Basis Extended Hanson--Koopman
CMH-17-1G provides Table 8.5.15, which contains factors for calculating
A-Basis values using the Extended Hanson--Koopman nonparametric method.
That table is reproduced in part here and the factors are compared with
those computed by `cmstatr`. More extensive checks are performed in the
unit test of the `cmstatr` package. The factors computed by `cmstatr` are
very similar to those published in CMH-17-1G.

```{r}
tribble(
  ~n, ~k,
  2, 80.0038,
  4, 9.49579,
  6, 5.57681,
  8, 4.25011,
  10, 3.57267,
  12, 3.1554,
  14, 2.86924,
  16, 2.65889,
  18, 2.4966,
  20, 2.36683,
  25, 2.131,
  30, 1.96975,
  35, 1.85088,
  40, 1.75868,
  45, 1.68449,
  50, 1.62313,
  60, 1.5267,
  70, 1.45352,
  80, 1.39549,
  90, 1.34796,
  100, 1.30806,
  120, 1.24425,
  140, 1.19491,
  160, 1.15519,
  180, 1.12226,
  200, 1.09434,
  225, 1.06471,
  250, 1.03952,
  275, 1.01773
) %>%
  rowwise() %>%
  mutate(z_calc = hk_ext_z(n, 1, n, 0.99, 0.95)) %>%
  mutate(abs_difference = expect_lt(abs(k - z_calc), 5e-4))
```

## Nonparametric B-Basis Binomial Rank
CMH-17-1G Table 8.5.12 provides factors for computing B-Basis values
using the nonparametric binomial rank method. Part of that table is
reproduced below and compared with the results of `cmstatr`.
The results of `cmstatr` are similar to the published values.
A more complete comparison is performed in the units tests of the
`cmstatr` package.

```{r}
tribble(
  ~n, ~rb,
  29, 1,
  46, 2,
  61, 3,
  76, 4,
  89, 5,
  103, 6,
  116, 7,
  129, 8,
  142, 9,
  154, 10,
  167, 11,
  179, 12,
  191, 13,
  203, 14
) %>%
  rowwise() %>%
  mutate(r_calc = nonpara_binomial_rank(n, 0.9, 0.95)) %>%
  mutate(test = expect_equal(rb, r_calc)) %>%
  select(-c(test))
```

## Nonparametric A-Basis Binomial Rank
CMH-17-1G Table 8.5.13 provides factors for computing B-Basis values
using the nonparametric binomial rank method. Part of that table is
reproduced below and compared with the results of `cmstatr`.
The results of `cmstatr` are similar to the published values.
A more complete comparison is performed in the units tests of the
`cmstatr` package.

```{r}
tribble(
  ~n, ~ra,
  299, 1,
  473, 2,
  628, 3,
  773, 4,
  913, 5
) %>%
  rowwise() %>%
  mutate(r_calc = nonpara_binomial_rank(n, 0.99, 0.95)) %>%
  mutate(test = expect_equal(ra, r_calc)) %>%
  select(-c(test))
```



# Session Info
This copy of this vignette was build on the following system.

```{r}
sessionInfo()
```

# References

