---
title: |
  cmstatr: An R Package for Statistical Analysis of Composite Material Data
tags:
  - R
  - statistics
  - composite materials
  - material science
authors:
  - name: Stefan Kloppenborg
    orcid: 0000-0002-1908-5214
    affiliation: 1
affiliations:
 - name: Comtek Advanced Structures, Ltd.
   index: 1
date: "3/22/2020"
bibliography: paper.bib
output:
  md_document:
    preserve_yaml: true
    variant: markdown
    pandoc_args: "--atx-headers"
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
Strength data for composite materials used in aerospace applications,
such as carbon fiber and fiberglass reinforced composites, are normally analyzed 
using statistical methods because of the 
inherent variability in the constituent materials and in the processing.
The design standards for civil aviation require that the probability
of structural failure due to this variability to be minimized, and to do
so, the designer must use what are called "Design Values" for each material 
in stress analyses and ensure they
exceed the actual stresses experienced by those materials in service. 
Design Values are determined such that, with $95\%$ confidence, they are 
either the $99\%$ or $90\%$ one-sided lower confidence bound of the material 
strength,
depending on the type of structure. These one-sided tolerance bounds
are referred to as A-Basis and B-Basis values, respectively. The statistical 
methods for calculating these 
basis values are outlined in Composite Materials Handbook, Volume 1, 
Revision G, or CMH-17-1G in short [@CMH171G].
The use of these methods is widely accepted by industry and civil 
aviation regulators.

Currently, many users use MS Excel spreadsheets to perform these
analyses. The MS Excel spreadsheets typically used, such as 
`STAT-17` [@STAT-17],
`ASAP` [@Raju_Tomblin_2008] and `CMH17-STATS` [@CMH17-STATS], use 
password-protected `VBA` macros to
perform the computations. As such, the code cannot be audited by 
the user. 
`cmstatr` is an R package that addresses this issue by implementing 
the same statistical analysis techniques 
found in CMH-17-1G in an open-source environment.


# Statement of Need
The purpose of `cmstatr` is to:

- Provide a consistent user interface for computing A- and B-Basis
  values and performing the related diagnostic tests in the `R`
  programming environment
- Allow auditing of the code used to compute A- and B-Basis values,
  and performing the related diagnostic tests
- Enable users to automate computation workflows or to perform
  simulation studies


# Implementation Goals
`cmstatr` aims give a consistent interface for the user. Most functions
are written to work with the `tidyverse` [@tidyverse] and most functions
have similar argument lists. The intent is to make the package easy to
learn and use.

The implementation of `cmstatr` also aims to avoid the use of look-up 
tables
that are prevalent in calculation spreadsheets and minimize the use of 
approximations. 
While this decision leads to increased computation time, the typically small 
data sets (tens to hundreds
of observations) associated with composite material test data, and the
speed of modern computers make this practical for interactive programming.

# Example Usage
Normally, to use `cmstatr` the user will load `cmstatr` itself as well as
the `tidyverse` package.

```{r message=FALSE}
library(cmstatr)
library(tidyverse)
```

`cmstatr` contains some example data sets, which can be used to demonstrate
the features of the package. One of those data sets --- `carbon.fabric.2` ---
will be used in the following example. This data set contains results from
several mechanical tests of a typical composite material, and contains the
typical measurements obtained from a test lab. In the following
examples, results from tension testing in the warp fiber direction (`WT`) 
will be used.
Part of this data set is shown below.

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  head(10)
```

One common task is to calculate B-Basis values. The
single-point basis functions perform the following tests:

- the maximum normed residual test for outliers within a batch [@CMH171G],
- the Anderson-Darling k-Sample test to check if batches are drawn from
  the same (unspecified) distribution [@Scholz_Stephens_1987], 
- the maximum normed residual test for outliers within the data, and 
- the Anderson-Darling test for a particular probability 
  distribution [@Lawless_1982].

Assuming that the data from the warp tension (WT) tested at 
elevated-temperature/wet condition (ETW) follows
a normal distribution, then this can be done using the function:

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  filter(condition == "ETW") %>%
  basis_normal(strength, batch)
```

All of the various basis functions perform diagnostic tests for each of the 
statistical tests mentioned above. If any
of the diagnostic tests failed, a warning is shown and the test failure
is also recorded in the returned object (and shown in that object's
`print` method). In the example above, the output shows that the
Anderson-Darling test for normality [@Lawless_1982] rejects the hypothesis
that the data is drawn from a normal distribution. 

Two non-parametric basis calculations, based on [@Guenther_1969] and
[@Vangel_1994] are also implemented in `cmstatr`. These functions 
perform the same
diagnostic tests, but omits the Anderson-Darling test for a particular
distribution.

The diagnostic test can be run directly using `cmstatr` as well.
For example, the failed diagnostic test above can be run as follows:

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  filter(condition == "ETW") %>%
  anderson_darling_normal(strength)
```

If the failure of a diagnostic test is decided to be acceptable,
the test result can be overridden to hide the warning in the basis function 
output:

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  filter(condition == "ETW") %>%
  basis_normal(strength, batch, override = c("anderson_darling_normal"))
```

`cmstatr` also provides functions for calculating basis values from
data pooled across multiple testing environments, as recommended by [@CMH171G]. 
There are two methods of pooling: both calculate a measure of global variation
from the data in all tested environmental conditions, 
then they calculate a basis value for each condition using the measure of
the global variation and the individual mean values of each 
condition. The two methods of pooling have different underlying assumptions
and hence the data must pass a different set diagnostic tests for each of
the two functions.

# Validation and Comparison With Existing Tools
Where possible, `cmstatr` has been verified against the examples given
in the articles in which the statistical methods were published. Unit
tests have been written so that this verification is re-checked routinely
to prevent unintended regressions. Agreement between `cmstatr` and the
examples in the original articles is within the expected numeric accuracy.

`cmstatr` has also been verified against existing software, such as
`STAT-17` [@STAT-17], `ASAP` [@Raju_Tomblin_2008] and
`CMH17-STATS` [@CMH17-STATS] using several example data sets.
Agreement between `cmstatr` and the other software is generally good,
but some results differ slightly, likely due to approximations used
in the software.
Comparison between `cmstatr` and the other software is performed
within various unit tests to guard against future regressions.

The tests are automatically run each time a change is made to the code
of `cmstatr` using a continuous integration service. Additionally,
`CRAN` runs `R CMD check` on each package routinely.


# Reproducibility
It is envisioned that many users of `cmstatr` will use it within an 
R Notebook or a Jupyter Notebook. It is further envisioned that this notebook
will be directly converted into the statistical analysis report. If this is
done, the reader of the statistical report will be able to verify all of the
detailed steps used in the statistical analysis.

# Acknowledgement
The author would like to thank Mr. Billy Cheng for his contributions to 
`cmstatr` and this paper. The author would also like to thank Comtek
Advanced Structures Ltd. for its support in developing and releasing
the `cmstatr` package.

# References
