---
title: |
  cmstatr: An R Package for Statistical Analysis of Composite Material Data
tags:
  - R
  - statistics
  - composite materials
  - material science
authors:
  - name: Stefan Kloppenborg
    orcid: 0000-0002-1908-5214
    affiliation: 1
affiliations:
 - name: Comtek Advanced Structures, Ltd.
   index: 1
date: "1/29/2020"
bibliography: paper.bib
output:
  pdf_document: default
  md_document:
    preserve_yaml: true
    variant: markdown
    pandoc_args: "--atx-headers"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
A number of statistical techniques are commonly used when analyzing
strength datat for composite materials used in aerospace applications,
such as carbon fiber and fiberglass.
Currently, many users use MS Excel spreadsheets for performing this
analysis. `cmstatr` is an R package that implements the statistical
analysis techniques commonly used for composite material strength data.

The design standards for civil aviation require that the probability
of structural failure due to material variability is minimized. To do
so, the desginer must select Design Values for each material and compare
those to the stresses experienced by those materials. These Design
Values are selected so that, with $95\%$ confience, the Design Value
is $99\%$ or $90\%$ lower confidnece bound of the material strength,
depending on the type of structure. These one-sided tolerance bounds
are referred to as A-Basis and B-Basis values, respectively. Computing
these A- and B-Basis values is the main problem that `cmstatr` addresses.

A set of statistical methods are described in a publication called
the Composites Materials Handbook, or CMH-17-1G [@CMH171G].
The use of these methods is widely accepted by industry and civil
aviation regulators. The methods described in CMH-17-1G are implemented
in `cmstatr`.

The MS Excel spreadsheets typically used, such as `STAT-17` [@STAT-17],
`ASAP` [@Raju_Tomblin_2008] and
`CMH-17 STATS` [INSERT CITATION], use password-protected `VBA` macros to
perform the
computations. As such, the code cannot be audited by the user. `cmstatr`
aims to address this by providing open-source code for performing
these computations.

# Implementation Goals
`cmstatr` aims give a consistent interface for the user. Most functions
are written to work with the `tidyverse` [@tidyverse] and most functions
have similar argument lists. The intent is to make the package easy to
learn and use.

The implementation of `cmstatr` also aims to avoid the use of loopup tables
and minimize the use of approximations. While this decision leads to
increased computation time, the typically small datasets (tens to hundreds
of observations) associated with composite material test data, and the
speed of modern computers make this practical for interactive programming.

# Example Usage
Normally, to use `cmstatr` the user will load `cmstatr` itself as well as
the `tidyverse` package.

```{r}
library(cmstatr)
library(tidyverse)
```

`cmstatr` contains some example datasets, which can be used to demonstrate
the features of the package. One of those datasets --- `carbon.fabric.2` ---
will be used in the following example. This dataset contains results from
several mechanical tests of a typical composite material. and contains the
typical measurements obtained from a test lab. In the following
examples, results from the "warp tension" (`WT`) test will be used. Part of
this dataset is shown below.

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  head(10)
```

One common task is to calculate B-Basis values. Depending on the distribution
of the data, this can be done using one of several functions. Assuming that
the data from the warp tension (WT) elevated-temperature wet (ETW) strength follows
a normal distribution, this can be done as follows:

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  filter(condition == "ETW") %>%
  basis_normal(strength, batch)
```

All of the various basis functions perform diagnostic tests. If any
of the diagnostic tests fail, a warning is emitted and the test failure
is also recorded in the returned object (and shown in that object's
`print` method). In the example above, the output shows that the
Anderson-Darling test for normality [@Lawless_1982] rejects the hypothesis
that the data is drawn from a normal distribution. The
single-point basis functions perform the following tests:
the maximum normed residual test for outliers within a batch [@CMH171G],
the Anderson-Darling k-Sample test to check if batches are drawn from
the same (unspecified) distribution [@Scholz_Stephens_1987], 
the maximum normed residual test for outliers within the data, and the
Anderson-Darling test for a particular distribution [@Lawless_1982].

Two non-parametric basis calculations, based on [@Guenther_1969] and
[@Vangel_1994] are implemented. These functions perform the same
diagnostic tests, but skip the Anderson-Darling test for a particular
distribution.

The diagnostic test can be run directly using `cmstatr` as well.
For example, the failed diagnostic test above can be run as follows:

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  filter(condition == "ETW") %>%
  anderson_darling_normal(strength)
```

If it is decided that the failure of the diagnostic test is acceptable,
the test can be overriden to avoid a warning from being emitted by
the basis function:

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  filter(condition == "ETW") %>%
  basis_normal(strength, batch, override = c("anderson_darling_normal"))
```

`cmstatr` provides functions for calculating basis values based on
data pooled accross environments, as recommended by [@CMH171G]. These
functions use the variance observed in different environmental conditions
in the computation, but acknowlege the different mean values under each
environmental condition.

Another common statistical technique is to determine if a sample is drawn
from a particular population. This is often used.....

`cmstatr` also provides functions for computing 

```{r}
carbon.fabric.2 %>%
  filter(test == "WT") %>%
  filter(condition == "RTD")  %>%
    equiv_mean_extremum(strength, n_sample = 8, alpha = 0.05)
```

# Comparison With Existing Tools

...excel tools...

...reproducibility...

...avoidance of lookup tables and approximations...

...example use...

...comparison with existing tools...

# Reproducibility
It is envisioned that many users of `cmstatr` will...


# References
